\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{csquotes}

\begin{document}
\twocolumn

\title{Monday Reading Assessment: Chapters 22-27 of \textit{Last Place on Earth}, Chapters 4-8 of \textit{Deep Survival}}
\author{Prof. Jordan C. Hanson}

\maketitle


\section{Chapter 22 - The Base at Framheim}

\begin{enumerate}
\item Roald Amundsen and team successfully laid depots all the way to 82 degrees South.  (a) How far is this from the South Pole? (b) What kind of safety margins did this leave? \\ \vspace{2cm}
\end{enumerate}

\section{Chapter 23 - Sledging with the Owner}

\begin{enumerate}
\item Robert Falcon Scott and company laid their main supply depot, One Ton Depot, at 79 degrees, 28.5 minutes South.  How much farther is this from the pole than that of Amundsen? \\ \vspace{2cm}
\item One Ton Depot had one tonne of supplies in it.  The depots of Roald Amundsen and company totalled three tons.  What does this imply about the comparative range of action for each party? \\ \vspace{2cm}
\end{enumerate}

\section{Chapter 24 - The Pole Seeker Prepares}

\begin{enumerate}
\item Discuss how the pumps aboard the \textit{Tera Nova} represented a \textit{single point of failure}, leading to additional risk for journeys between New Zealand and Ross Island. \\ \vspace{2cm}
\item Note that, when Prestrud was computing copies of navigation tables, he had to perform some astronomical calculations and measure star positions.  This is only possible during Antarctic winter, when there is no sun.  What happened to the only Almanac that they brought?  The Almanac was used in conjunction with the astronomical calculations to check longitude and lattitude.  Does this represent a single point of failure? \\ \vspace{2cm}
\item How else was this party to measure distance and direction on the sledges in the field? \\ \vspace{2cm}
\item Discuss ways in which good leadership maintained the group dynamic over the winter in the Bay of Whales camp. \\ \vspace{2cm}
\end{enumerate}

\section{Chapter 25 - Wintering at Cape Evans}

\begin{enumerate}
\item What decisions did Scott make about \textit{transport} to and from the South Pole? \\ \vspace{2cm}
\item Describe the group dynamic at the Cape Evans camp. \\ \vspace{2cm}
\end{enumerate}

\section{Chapter 26 - False Start}

\begin{enumerate}
\item What was Roald Amundsen's main mistake in selecting the start date for the journey to the South Pole? \\ \vspace{1cm}
\end{enumerate}

\section{Chapter 27 - Scott's Caravan}

\begin{enumerate}
\item How many types of transport did Scott use on the caravan to the South Pole?  What happened to each? \\ \vspace{2cm}
\end{enumerate}

\section{(Open Discussion on Deep Survival)}

\end{document}
