\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{csquotes}

\begin{document}

\title{Activity: Large Creatures in Antarctica}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Bears in Arctic and Antarctica}

\begin{enumerate}
\item How many individuals remain in the Spirit Bear population?  Where do they live? \\ \vspace{1cm}
\item Name three of the animals in the photo essay you recognize. \\ \vspace{1cm}
\item What do polar bears eat?  Name specific species.  How many pounds of meat was the polar bear in the essay consuming?  \\ \vspace{2cm}
\item Given current projections, in how many years (according to the essay) will polar bears go extinct? \\ \vspace{1cm}
\end{enumerate}

\section{Multi-year ice}

\begin{enumerate}
\item According to the essay, what is the function of the multi-year ice as it pertains to the microbiology of the ecosystem? \\ \vspace{2cm}
\item What types of organisms feed off of the underside of the ice? \\ \vspace{1cm}
\item As the food chain builds up, it eventually supports huge creatures like bowhead whales.  How long can bowheads live?  What was going on in history that long ago? \\ \vspace{2cm}
\end{enumerate}

\section{Leopard Seals}

\begin{enumerate}
\item What do leopard seals eat?  How large are they? \\ \vspace{1cm}
\item Describe some of the animal behaviors the explorer encounters in the water with the seals. \\ \vspace{2cm}
\end{enumerate}

\section{Orcas and Hunting Technique}

\begin{enumerate}
\item What happens to the ice under the photographer's feet when he saw the orcas?  This turned out to be a warning about the ``state of the ice'' in the region of Terra Nova's landing. \\ \vspace{1cm}
\item Do you recall what happens next, to the motor sledge, given the thinness of the ice? \\ \vspace{1cm}
\item Describe the hunting technique of the orcas for seal.  What does this tell us about their social relationships?  \\ \vspace{1cm}
\end{enumerate}

\end{document}
