\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{csquotes}

\begin{document}

\title{Activity on Leadership: The West Point Lecture on Leadership}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Introduction to Part 1}

The last time we were together, we listened and read the first half of a speech given by a Dr. William Deresiewicz, a professor of English from Yale University, given to the \textit{freshmen} (plebians) of West Point Military Academy.  The title of the lecture was \textbf{Solitude and Leadership}.  Let's recall some of the ideas from the lecture and connect them to the events of the final chapters of \textit{The Last Place on Earth.}

\section{What a Leader is Not}

\begin{enumerate}
\item The author of the speech recalls being on the admissions committee for Yale University.  After reviewing applications, the committee began to look at the test scores, grades, and high school rankings of the applicants.  But it turned out this was not enough to make the final decisions.  They looked at extra-curricular activities and found that students were listing 10 to 12 of them in hopes of impressing the committee.  Students later told the professor they felt like ``excellent sheep,'' or ``world-class hoop jumpers.''  As a group, discuss some examples of extra-curricular activities you do.  Honestly assess whether or not they are helping to educate you or to mold you into a leader. Write down some of the main discussion points here.  \\ \vspace{3cm}
\item The author goes on to relate the idea of \textit{hierarchy} to bureaucracy in our society, and to The Company in the novel \textit{The Heart of Darkness} by Joseph Conrad.  Some people view a leader as someone who rises to the top of a hierarchy.  Discuss how this idea applies to the life of Captain Robert Falcon Scott.  Describe the type of hierarchy to which Captain Scott aspired.  What rank did he hope to achieve, and why?  (Write down the main discussion points here).  \\ \vspace{3cm}
\item Apply the same thought to Roald Amundsen.  Now that you are learning about the end of his life, look back across his life and answer the same question: is there a hierarchy to which Roald Amundsen ascribes?  Is there a sense of true leadership that you learn from his life, and if so what specific examples inform your opinion?  (Write down the main discussion points here). \\ \vspace{3cm}
\item Recall the description of The Big Boss of the Central Station in the Heart of Darkness from part I of the speech.  What qualities does this character lack that we usually attribute to a true leader? \\ \vspace{2cm}
\end{enumerate}

\section{Part II of the Lecture}

(Listen carefully and take notes as the author continues to shape the idea of a leader). \\ \vspace{1cm}

\section{What a Leader Is}

\begin{enumerate}
\item What does the author mean by writing that a good leader is a good \textit{thinker?}  In what ways does the person he chose as an example, General David Petraeus, reach this goal of leadership through being an independent thinker?  (Write down the main discussion points here). \\ \vspace{2.5cm}
\item Having heard the author's words, define the following concepts and discuss how they are related: \textit{concentration}, \textit{solitude}, and \textit{introspection.} (Write down the main discussion points here). \\ \vspace{3cm}
\item Why, according to the author, is reading a carefully written book better than learning through social media? How does reading a book relate to solitude?  (Write down the main discussion points here).  \\ \vspace{3cm}
\item \textbf{The idea of Friends is central to Whittier College.}  What does the author write about the value of \textit{true} friends as it relates to solitude?  How do friends help us find ourselves? (Write down the main discussion points here). \\ \vspace{3cm}
\end{enumerate}

\end{document}
