\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{csquotes}

\begin{document}

\title{Thursday Reading Assessment: Chapters 18-20 of \textit{Last Place on Earth}}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Memory Bank}

\begin{itemize}
\item $s = R\theta$ Using radius of the Earth $R$ and latitude angle $\theta$ to calculate a distance.
\item $R_{E} = 6371$ km (radius of the Earth).
\item Convert an angle in degrees to radians: $\theta_{rad} = \theta^{\circ} \frac{\pi}{180^{\circ}}$.
\end{itemize}

\section{Chapter 18 - Volte-Face}

\begin{enumerate}
\item Roald Amundsen chose to land and establish winter quarters not on firm land, but on the \textit{Ross Ice Shelf itself.}  What were some of the reasons he chose to do this? \\ \vspace{1.5cm}
\item The landing point of Amundsen and his companions was about 1.0 degree of latitude closer to the pole than Robert Falcon Scott.  What distance is this in kilometers?  Use the formulas in the memory bank to work it out. \\ \vspace{1.5cm}
\item Roald Amundsen studied the work of Ernest Shackleton in his attempt on the South Pole on the \textit{Nimrod} expedition.  What conclusions did he draw, and what did he decide to do with margins of error? \\ \vspace{3cm}
\end{enumerate}

\section{Chapter 19 - On the Terra Nova}

\begin{enumerate}
\item Part of Robert Falcon Scott's plan for tackling the surface travel from McMurdo sound to the South Pole was the \textit{motor sledge.}  Describe the scene when they first tested the complete prototype near Christiania.  What problem occurred, and how did they remedy the situation?  What was to be learned from this situation? \\ \vspace{3cm}
\item Although Scott tried to meet with Amundsen to learn new things about the Antarctic, Amundsen managed to avoid Scott while Scott was in Christiania.  Why did Amundsen do this? \\ \vspace{2cm}
\end{enumerate}

\section{Chapter 20 - The Secret Disclosed}
\begin{enumerate}
\item Describe the scene when Amundsen boarded the \textit{Fram} and left for Antarctica.  From where did they leave?  What items did they load on board?  Where was Fridtjof Nansen? \\ \vspace{1.5cm}
\item Once on board the \textit{Fram}, and out to sea, Roald Amundsen began to type letters relating his secret plan to change course for the South Pole.  Who were some of the people to whom he wrote letters?  \\ \vspace{1.5cm}
\item Describe the scene as Amundsen revealed his secret plan to reach the South Pole to his shipmates.  How did they respond?  What does this tell us about his leadership style?  \\ \vspace{1.5cm}
\item There is an interesting detail about \textit{wireless} (they mean radio communications).  What about the race to the pole between Scott and Amundsen, and the press, would have changed if Scott had set up wireless communications to New Zealand?  (Recall the writing about ``the race to the cable head''). \\ \vspace{2.5cm}
\item The crew had a polar exploration library on board the \textit{Fram}, including accounts of the expeditions of \textit{Discovery} and \textit{Nimrod}.  There were several reasons for having these books on board.  What were some of them?
\end{enumerate}
\end{document}
