\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}

\title{Thursday Reading Assessment: Chapters 5-8 of \textit{Last Place on Earth}}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Chapter 5: A Sailor On Skis}

\begin{enumerate}
\item The author writes in this chapter about how Roald Amundsen and his brother Gustav, along with various other companions, begin to explore Hardangervidda on skis.  What are some of the survival skills they learn during these treks?  What risks do they encounter, and why was it important for Roald to learn them at the time? \\ \vspace{4cm}
\item What was the purpose of the Artic ships on which Roald Amundsen serves in this chapter?  What were they harvesting up North? \\ \vspace{2cm}
\end{enumerate}

\section{Chapter 6: Into the Antarctic Night}

\begin{enumerate}
\item Captain De Gerlache of the \textit{Belgica} and crew, including Amundsen, are exploring Graham Land of Antarctica in this chapter.  This is the ``tail'' of Antarctica, that peninsula that sticks out towards Tierra del Fuego at the southern tip of Latin America.  They explore De Gerlache strait, and as the sun sets for the winter the ocean around them begins to freeze.  Why does this worry many in the crew? \\ \vspace{3cm}
\item What is scurvy, and how did Captain De Gerlache and company prevent themselves from dying of scurvy?  \textbf{Bonus point: which person on the voyage first taught the group this trick for defeating scurvy?}\\ \vspace{3cm}
\end{enumerate}

\section{Chapter 7: Amundsen's First Command}

\begin{enumerate}
\item Roald Amundsen wants to explore the North West Passage, and be the first person to command a ship through it in one go.  However, there had to be a scientific reason for the expedition: to locate the geomagnetic north pole.  Why is this location distinct from the \textit{geographic} north pole, and why does it constantly have to be re-located? \\ \vspace{3cm}
\end{enumerate}

\section{Chapter 8: Artic Apprentiship}

\begin{enumerate}
\item As the \textit{Gj{\o}a} explored Northern Canada in the North West Passage voyage, the crew search for the direction towards the \textit{geomagnetic} North Pole.  Why do they need a special instrument for this?  \textit{Hint: a regular compass does not work in this case}. \\ \vspace{3cm}
\item During the North West Passage voyage of the Gj{\o}a, Amundsen and company encounter members of the Netsilik community and are greeted ``as old friends,'' even though the Netsilik had never seen Europeans before.  Amundsen ventures to the tribe's igloo village \textit{alone,} and begins to learn from them.  Name as many techniques as you can that Amundsen learned from the Netsilik.
\end{enumerate}

\end{document}
