\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}
\twocolumn

\title{Wednesday Warm Up: Navigation, Latititude, and Longitude}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Memory Bank}

\begin{enumerate}
\item $s = R\theta$ ... Distance $s$, given Earth radius $R$, and change in \textit{latitude} $\theta$.
\item $s = \phi R\cos\theta$ ... Distance $s$, given Earth radius $R$, a fixed \textit{latitude} $\theta$, and change in \textit{longitude} $\phi$.
\item $v = \Delta x/\Delta t$ ... The speed, $v$, given the distance $\Delta x$ and the time duration $\Delta t$.
\item 1 knot = 1 nautical mile per hour
\item 1 nautical mile = distance corresponding to 1 minute of latitude
\item 1 degree of latitude or longitude equals 60 minutes
\item 1 minute of latitude or longitude equals 60 seconds
\item Pythagorean theorem: $\sqrt{x^2 + y^2} = h$, where $x$ is the horizontal distance, $y$ is the vertical distance, and $h$ is the distance travelled.
\end{enumerate}

\section{Speed, Distance, Time}

\begin{enumerate}
\item Suppose you hike South for 10 km.  If the hike takes 4 hours, what was your average speed? \\ \vspace{2cm}
\item Suppose you hike North for 3 hours.  If your average speed is 1.5 km per hour, how far did you travel? \\ \vspace{2cm}
\item Suppose you hike West for 7 miles at an average pace of 3.5 miles per hour.  How long does this take? \\ \vspace{2cm}
\item Suppose you hike South for 3 km, then West for 4 km.  (a) How far are you from the original location? (b) If your average speed is 2 km per hour, how long does the hike take? \\ \vspace{2cm}
\end{enumerate}

\section{Navigation}

\begin{enumerate}
\item Suppose we sail North at 12 knots for 24 hours.  (a) How far have we travelled? (b) What is our change in latitude? \\ \vspace{4cm}
\item Suppose we sail East at 10 knots for 15 hours, and our latitude is 45 deg North. (a) How far have we travelled? (b) What is our change in longitude? \\ \vspace{4cm}
\end{enumerate}

\end{document}
