\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}
\twocolumn

\title{Reading Quiz: Chapters 5-7 of \textit{Last Place on Earth}, and ``Cape Foulweather'' of \textit{Horizon}}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Chapter 5: A Sailor on Skis}

\begin{enumerate}
\item List at least three ways Roald Amundsen began to learn about polar exploration. \\ \vspace{2cm}
\item How did Roald react to sealing to harvest blubber and seal fur? \\ \vspace{2cm}
\item What physical ``defect'' did Roald wish to conceal during a medical examination prior to his mandatory military service? Does this reveal something about his character? \\ \vspace{2cm}
\item Though Roald and his brother Leon crossed Hardangervidda successfully, it was dangerous.  (a) What improvements did they make relative to the previous attempt?  (b) What went wrong?\footnote{Hint: was it weather, compasses, snow types, lack of food ... ?}\\ \vspace{2cm}
\end{enumerate}

\section{Chapter 6: Into the Antarctic Night}

\begin{enumerate}
\item \textbf{Nansen made it!}  Describe how Fridtjof Nansen led the polar expedition towards the North Pole.  (a) How close did they come to reaching the North Pole? (b) How did they achieve this? (c) How did they make it back to civilization? \\ \vspace{2cm}
\item As Amundsen and De Gerlache prepared for their first Antarctic expedition with the \textit{Belgica} in 1897, note some of the ways they prepared themselves mentally.  For example, what new languages did they have to learn?  What intellectual skills did they need to polish? \\ \vspace{2cm}
\item (a) When the \textit{Belgica} entered Graham Land, into the Gerlache Strait, what kind of observations were recorded by Amundsen? (b) Once the \textit{Belgica} froze within pack ice for the winter, what mental and physical challenges were faced by the crew? \\ \vspace{1.5cm}
\item How did the crew survive scurvy? \\ \vspace{1.5cm}
\item After a narrow escape from the pack ice in March, 1899, the \textit{Belgica} made it to Punta Arenas.  Summarize the discoveries of the expedition.  What historic firsts were achieved?
\end{enumerate}

\end{document}
