\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}
\twocolumn

\title{Reading Quiz: Chapters 1-4 of \textit{Last Place on Earth}}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Chapter 1: Rivals for the Pole}

\begin{enumerate}
\item In what year did Roald Amundsen and Robert Falcon Scott (and company) set out for the South Pole?
\begin{itemize}
\item A: 1891
\item B: 1901
\item C: 1911
\item D: 1921
\end{itemize}
\item From what country was Roald Amundsen and most of his colleagues?
\begin{itemize}
\item A: Sweden
\item B: England
\item C: Norway
\item D: Denmark
\end{itemize}
\end{enumerate}

\section{Chapter 2: Forerunners}

\begin{enumerate}
\item In 1768-9, a reknowned British explorer named Captain James Cook sailed to Antarctica by way of Tahiti, along with an astronomer named Charles Green.  What was the \textit{public} reason for the expedition?  What was the \textit{secret} reason for the expedition, that prompted them to sail to Antarctica in addition to Tahiti? \\ \vspace{3cm}
\item Lieutenant James Clark Ross was a naval officer who was ordered by the Admiralty to complete a ``magnetic survey'' of Earth's southern lattitudes.  When he learned that American and French explorers could reach the \textit{magnetic} South Pole first, he decided to instead explore uncharted Antarctic waters.  What are some of his expedition's major discoveries?  What geological feature is named after James Clark Ross?  \\ \vspace{3cm}
\item Consider the following quote: ``While Franklin and his men were dying of hunger, Eskimos around them were living off the land in comparative plenty.  But Franklin was hampered by grotesquely unsuitable methods, the product of rigid thought and incapacity to adapt to circumstances.''  What passages or events from ``Looking for a Ship'' in \textit{Horizon}, by Barry Lopez, connect to the notion that we have much to gain by learning from those who are different from us? \\ \vspace{5cm}
\end{enumerate}

\section{Chapter 3: The Last of The Vikings}
\begin{enumerate}
\item Circle all of the following activities in which the Amundsen family engaged as young people:
\begin{itemize}
\item Sailing
\item Fur trading, trapping
\item Skiing
\item Doing physics
\item Farming
\item Shipbuilding
\item Playing baseball
\item Studying economics
\end{itemize}
\item Consider Roald Amundsen's early experience with skiing in Northern Norway.  (a) How has skiing evolved over time? (b) What kind of wax was used to enable wooden ski to slide on snow?  Did it work well at the time? \\ \vspace{3cm}
\end{enumerate}

\section{Chapter 4: The Nansen Spirit}
\begin{enumerate}
\item Describe briefly Fridtjof Nansen's introduction to the global stage.  What was his first major accomplishment?  Why was it important to Norway? \\ \vspace{3cm}
\item From whom did Fridtjof Nansen learn to make proper polar clothing (for the time)?
\begin{itemize}
\item The Maori
\item The Godthaab Inuits
\item English naval officers
\item American explorers
\end{itemize}
\item Who is Askeladden? \\ \vspace{2cm}
\item List some examples of how Nansen and Amundsen encountered and built leadership characteristics. \\ \vspace{4cm}
\end{enumerate}

\section{Looking for a Ship}
\begin{enumerate}
\item Describe the significance of painting to the author.  What emotions and ideas are evoked for the author by the painting \textit{Remember?} \\ \vspace{4cm}
\item Why does the author feel it is so important to travel and interact with people from far away lands? \\ \vspace{4cm}
\item The author ends the essay by writing about love.  Who do you love?  Would you be willing to travel far away from your homeland, far from your loved ones?  Perhaps you have already tried this.  What have you learned in a journey that you have taken?
\end{enumerate}

\end{document}
