\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}
\twocolumn

\title{Reading Quiz: ``Cape Foulweather'' of \textit{Horizon}, and Chapter 8 of \textit{The Last Place on Earth}}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{\textit{Horizon}: Cape Foulweather}

\begin{enumerate}
\item \textbf{Topic for group discussion.} There are a great number of deep, humanistic themes in this essay.  These include, but are not limited to: colonialism, disappearing cultures, local and global worldviews, diverse ``ways of knowing,'' treating human suffering, and cross-cultural interactions.  Choose a theme from the essay that you would like to discuss with your class mates. When we're done completing the reading quiz, we can reflect on these as a group.
\item Where is Cape Foulweather, and what is its historical significance? \\ \vspace{1cm}
\item (a) Who lived there when Cpt. Cook arrived as an explorer? (b) What were the names of the ships of the expedition of Cpt. Cook that year? \\ \vspace{1cm}
\item Describe the significance of the geographic exploration of Cpt. Cook for scientific and human progress? \\ \vspace{1cm}
\item What are some examples of indigenous knowledge that was shared with the explorers? \\ \vspace{2cm}
\item (a) What feat of exploration was achieved by Don Walsh and Jacques Piccard? (b) Where is a place you have traveled that has given you a sense of \textit{wonder?} \\ \vspace{2cm}
\item In the essay, there is an interlude describing the astrophysics learned from the Hubble Telescope.  The author relates his experiences with telescopes to encountering a ``piece of the solar system'' in the Trans-Antarctic mountains.  What did he and his companions find? \\ \vspace{2cm}
\item Who was Ranald MacDonald, and what cultural interaction did he initiate? What behaviors did he demonstrate that aided his efforts? \\ \vspace{2cm}
\item Consider the following quote ``Whether the change facing a people comes on swiftly ... or slowly ... the responsibility of the wisdom keeper is to recognize the early signs of significant change, to look into the past, and locate, again, a through line to the future.'' Apply this idea to a problem facing our community today. \\ \vspace{3cm}
\end{enumerate}

\section{Chapter 8: Arctic Apprentiship}

\begin{enumerate}
\item List five polar survival and travel techniques that the \textit{Gjoa} crew learned from the Netsilik.
\end{enumerate}

\end{document}
